---
output: rmarkdown::github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r opts, echo = FALSE}
knitr::opts_chunk$set(
                      collapse = TRUE,
                      warning = TRUE,
                      message = TRUE,
                      width = 120,
                      comment = "#>"
                      )
```
[![Build Status](https://travis-ci.org/osm-router/osmprob.svg?branch=master)](https://travis-ci.org/osm-router/osmprob)
[![Build status](https://ci.appveyor.com/api/projects/status/lw5a4udgpjpaf2if?svg=true)](https://ci.appveyor.com/project/karpfen/osmprob)
[![codecov](https://codecov.io/gh/osm-router/osmprob/branch/master/graph/badge.svg)](https://codecov.io/gh/osm-router/osmprob)
[![Project Status: WIP - Initial development is in progress, but there has not yet been a stable, usable release suitable for the
public.](http://www.repostatus.org/badges/latest/wip.svg)](http://www.repostatus.org/#wip)
[![CRAN\_Status\_Badge](http://www.r-pkg.org/badges/version/osmprob)](http://cran.r-project.org/web/packages/osmprob)

# osmprob

R package for probabilistic OpenStreetMap routing

------

## Install

```{r, eval=FALSE}
devtools::install_github ('osm-router/osmprob')
```

```{r, echo=FALSE, eval=FALSE}
devtools::load_all ('.', export_all=FALSE)
devtools::document ('.')
devtools::check ('.')
```

## Test road graph

A sample graph with OSM data can be found in `tests/osm-ways-munich`. It was downloaded and saved with
[osmdata](https://github.com/osmdatar/osmdata) like this:

```{r eval=FALSE}
q <- osmdata::opq (bbox = c (11.58, 48.14, 11.585, 48.145))
q <- osmdata::add_feature (q, key = 'highway')
dat <- osmdata::osmdata_xml (q, "osm-ways-munich.osm")
```

To prepare the graph for routing, all non-accessible edges are removed using `makeCompactGraph ()`, so only the largest connected part of the
original graph remains. The resulting graph is stored as a `data.frame`. A sample can be found in `tests/compact-ways-munich.Rda`.

After the preprocessing is done, `plotGraph ()` can be used to display the compact sample graph in a `shiny`/`leaflet` web app.

### From sample data

To load, process and display a sample graph, run
```{r eval=FALSE}
devtools::load_all (export_all = FALSE)
library (magrittr)
graph <- readRDS ("tests/compact-ways-munich.Rda") %>% makeCompactGraph
#startPt <- graph$compact$from_id [1]
#endPt <- graph$compact$to_id [600]
startPt <- graph$compact$from_id [350] %>% as.character %>% as.numeric
endPt <- graph$compact$to_id [615] %>% as.character %>% as.numeric
path <- getShortestPath (graph, startPt, endPt)
prob <- getProbability (graph, startPt, endPt, eta = 1)
plotMap (prob, path)
```

### From OSM data

```{r eval=FALSE}
devtools::load_all (export_all = FALSE)
start_pt <- c (11.58, 48.14)
end_pt <- c (11.585, 48.145)
graph <- download_graph (start_pt, end_pt)
route_start <- graph$compact$from_id [1]
route_end <- graph$compact$to_id [600]
path <- getShortestPath (graph, route_start, route_end)
prob <- getProbability (graph, route_start, route_end, eta = 1)
plotMap (prob, path)
```

## Code of Conduct

Please note that this project is released with a [Contributor Code of Conduct](https://github.com/osm-router/osmprob/blob/master/CONDUCT.md). By participating in this project you agree to abide by its terms.
